---
title: "FFT"
output: 
  slidy_presentation: 
    css: ~/Documents/Projects/Packages/represtools/vignettes/My.css
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
```

#

> "You say you want a convolution? Well, you know we all want to change the world."

A convolution is merely the sum of a set of numbers

# Pascal's triangle

This is just a continuous sum of a vector which grows by one element at each iteration.

```{r }
pascalLoop <- function(n){
  x <- numeric(n)
  n_1 <- n - 1
  x <- c(1, rep(0, n_1))
  for (i in seq.int(n_1)){
    x <- c(x[1], tail(x, n_1) + head(x, n_1))
  }
  
  x
}

pascalLoop(4)
```

# Pascal's triangle with FFT

```{r }
pascalFFT <- function(n){
  x <- c(1, 1, rep(0, n-2)) 
  x <- fft(x)
  x <- x ^ (n-1)
  x <- fft(x, inverse = TRUE)
  x <- round(Re(x / n) )

  x
}

pascalFFT(4)
```

```{r echo=FALSE, eval=FALSE}
identical(pascalLoop(4), pascalFFT(4))
head(pascalFFT(55))
head(pascalLoop(60))
identical(pascalLoop(60), pascalFFT(60))
```

# Shift development patterns

We can also do this

```{r }
QuarterWeights <- c(0.25, 0.25, 0.25, 0.25, 0, 0, 0, 0)
AccidentQuarter <- c(0, 0.4, 0.4, 0.2, 0, 0, 0, 0)
AccidentYear <- fft(fft(AccidentQuarter) * fft(QuarterWeights), inverse = TRUE) %>% Re()
AccidentYear <- AccidentYear / 8
round(AccidentYear, 4)
```

# Aggregate loss distributions

* FFT is great for generating aggregate loss distributions. 
* Adding probability is a lot like adding integers. Rather than raising to a power, we use a probability generating function.
* FFT is a discrete approximation, not a closed form solution. Careful with how many values you use!

# How do we do it?

1. Form a vector with the severity probabilities
2. Take the Fourier transform
3. Apply the probability generating function
4. Take the inverse Fourier transform

# Klugman

```{r}
PoissonPGF <- function(x, lambda){
  x <- lambda * (x - 1)
  x <- exp(x)
  x
}

FormS <- function(px, n){
  px <- c(px, rep(0, n-4))
  px <- fft(px)
  px <- PoissonPGF(px, 3)
  s <- Re(fft(px, inverse=TRUE) / n)
}

px <- c(0, 0.5, 0.4, 0.1)

aggs <- lapply(c(8, 4096), FormS, px = px)
```

# Cyclical overflow

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Set = c(rep(8, 8), rep(4096, 4096)), x = c(1:8, 1:4096), y = unlist(aggs))
plt <- ggplot(data=filter(df, x<=30), aes(x=x, y=y, color=as.factor(Set))) + geom_point()
plt <- plt + guides(color = guide_legend(title = NULL))
plt
```

# A circle has no end

```{r echo=FALSE}
RootsOfUnity <- function(n){
  k <- seq(0, n-1)
  exponent <- complex(imaginary = 2 * pi * k / n)
  w <- exp(exponent)
  w
}

myRoots <- lapply(c(8, 4096), RootsOfUnity)

df1 <- data.frame(Real = Re(myRoots[[1]]), Imaginary = Im(myRoots[[1]]))
df1$Set <- 8

df2 <- data.frame(Real = Re(myRoots[[2]]), Imaginary = Im(myRoots[[2]]))
df2$Set <- 4096

df <- rbind(df1, df2)
df$Alpha <- 1/ df$Set

plt <- ggplot(df, aes(x = Real, y = Imaginary, color=Set, alpha = 1/Set)) + geom_point()
plt + theme(legend.position = "none")
```

```{r, eval=FALSE, include=FALSE}
x <- c(1,0,0,0,0)
x1 <- fft(x)
x1 <- x1^6
x1 <- fft(x1, inverse = TRUE) %>% Re() %>% round()
x1 / 5
```

```{r, eval=FALSE, include=FALSE}
x <- c(1, 0, 0, 0)
x <- c(x[1], tail(x, 3) + head(x, 3))
x <- c(x[1], tail(x, 3) + head(x, 3))
x <- c(x[1], tail(x, 3) + head(x, 3))
```

